QDNAseq segmentation

### 2019-10-30 QDNAseq segmentation.Rmd
updating to process downsampled

```{r}
m = melt(do.call(bind_rows, train.dists.by.kb), id.vars='kb', measure.vars=c('median'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0.95,1.05) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Training data median per QDNAseq bin size', y='median',x='') + theme_minimal()

m = melt(do.call(bind_rows, train.dists.by.kb), id.vars='kb', measure.vars=c('stdev'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0,1) +
  scale_fill_brewer(palette='Set1') + 
  labs(title='Training data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()

@@

```{r}
m = melt(do.call(bind_rows, val.dists.by.kb), id.vars='kb', measure.vars=c('median'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0.95,1.05) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation data median per QDNAseq bin size', y='median',x='') + theme_minimal()

m = melt(do.call(bind_rows, val.dists.by.kb), id.vars='kb', measure.vars=c('stdev'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0,1) +
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()

@@

```{r}
all = bind_rows(do.call(bind_rows, train.dists.by.kb) %>% mutate(cohort='training'),
                do.call(bind_rows, val.dists.by.kb) %>% mutate(cohort = 'validation'))
all = all %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))

m = melt(all, id.vars=c('kb','cohort'), measure.vars='stdev')
ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + ylim(0,1) +
  geom_jitter(width=0.2,alpha=0.5) + geom_violin(aes(fill=kb), alpha=0.5, show.legend=F) + 
  #geom_boxplot(aes(fill=kb), outlier.color=NA, show.legend = F) 
  scale_fill_brewer(palette='Set1') + labs(title='Data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()

m = melt(all, id.vars=c('kb','cohort'), measure.vars='median')
ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + 
  geom_jitter(width=0.2,alpha=0.5) + geom_violin(aes(fill=kb), alpha=0.5, show.legend=F) + ylim(0.95,1.1) +
  scale_fill_brewer(palette='Set1') + labs(title='Median per QDNAseq bin size', y='median',x='') + theme_minimal()

ggplot(all %>% filter(kb == '100kb'), aes(stdev)) + facet_grid(~cohort) + 
  geom_density(fill='lightblue',kernel='gaussian') + xlim(0,1)

@@

```{r}
plist = lapply(train.resids, function(df) {
  lim = c(0, max(df$varMAD_median))
  if (max(df$varMAD_median) > 0.05) lim = c(min(df$varMAD_median), 0.05)
  ggplot(df, aes(varMAD_median)) + geom_histogram(color='lightblue') + labs(title=unique(df$kb)) + xlim(lim)
})
do.call(grid.arrange, c(plist))

@@

m = melt(all.resids, id.vars=c('kb','cohort'), measure.vars = 'varMAD_median')
ggplot(m %>% filter(value < 0.1), aes(cohort, value)) + facet_grid(~kb) + 
  geom_jitter(width=0.2, alpha=0.5) + geom_violin(aes(fill=cohort), alpha=0.5, show.legend=F) + 
  labs(title='Post-segmentation MAD residuals', x='median MAD variance')

cutoffs = list('100kb' = 0.009, '15kb' = 0.011, '500kb' = 0.005, '50kb' = 0.011)

sapply(names(train.resids), function(kb) nrow( train.resids[[kb]] %>% filter(varMAD_median >= cutoffs[kb])))
sapply(names(val.resids), function(kb) nrow( val.resids[[kb]] %>% filter(varMAD_median >= cutoffs[kb])))

ggplot(do.call(bind_rows,train.resids), aes(kb, varMAD_median)) + geom_boxplot() + ylim(0,0.05)
```

@@

ggplot(m, aes(variable, value, color=cohort, group=cohort)) + ylim(0.95,1.05) + facet_grid(~kb) +  geom_jitter(position = position_dodge(width=1)) + geom_violin(position = position_dodge(width=1), alpha=0.5)

@@

```{r}
z.sds = lapply(train.tiles, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
})

>>>>

```{r}
m = melt(do.call(bind_rows, train.dists.by.kb), id.vars='kb', measure.vars=c('median'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) + ylim(0.95,1.05) + 
  geom_jitter(width=0.1) + 
  geom_violin(aes(fill=kb), show.legend = F) +
  geom_boxplot(width=0.1,alpha=0.5) +
  scale_fill_brewer(palette='Set1') + 
  labs(title='Training data median per QDNAseq bin size', y='median',x='') + theme_minimal()

m = melt(do.call(bind_rows, train.dists.by.kb), id.vars='kb', measure.vars=c('stdev'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) + ylim(0,1) +
  geom_jitter(width=0.1) + 
  geom_violin(aes(fill=kb), show.legend = F) +
  geom_boxplot(width=0.1,alpha=0.5) +
  scale_fill_brewer(palette='Set1') + 
  labs(title='Training data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()

@@

```{r}
m = melt(do.call(bind_rows, val.dists.by.kb), id.vars='kb', measure.vars=c('median'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) + ylim(0.95,1.05) + 
  geom_jitter(width=0.1) + 
  geom_violin(aes(fill=kb), show.legend = F) +
  geom_boxplot(width=0.1,alpha=0.5) +
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation data median per QDNAseq bin size', y='median',x='') + theme_minimal()

m = melt(do.call(bind_rows, val.dists.by.kb), id.vars='kb', measure.vars=c('stdev'))
m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
ggplot(m, aes(kb, value, group=kb)) +  ylim(0,1) +
  geom_jitter(width=0.1) + 
  geom_violin(aes(fill=kb), show.legend = F) +
  geom_boxplot(width=0.1,alpha=0.5) +
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()

@@

```{r, fig.width=10, fig.height=8}
all = bind_rows(do.call(bind_rows, train.dists.by.kb) %>% mutate(cohort='training'),
                do.call(bind_rows, val.dists.by.kb) %>% mutate(cohort = 'validation'))
all = all %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))

m = melt(all, id.vars=c('kb','cohort'), measure.vars='median')
ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + ylim(0.95,1.1) +
  geom_jitter(width=0.1) + 
  geom_violin(aes(fill=kb), show.legend = F) +
  geom_boxplot(width=0.1,alpha=0.5) +
  scale_fill_brewer(palette='Set1') + labs(title='Median per QDNAseq bin size', y='median',x='') + theme_minimal()

m = melt(all, id.vars=c('kb','cohort'), measure.vars='stdev')
ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + ylim(0,1) +
  geom_jitter(width=0.1) + 
  geom_violin(aes(fill=kb), show.legend = F) +
  geom_boxplot(width=0.1,alpha=0.5) +
  scale_fill_brewer(palette='Set1') + labs(title='Data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()

#ggplot(all %>% filter(kb == '100kb'), aes(stdev)) + facet_grid(~cohort) + 
#  geom_density(fill='lightblue',kernel='gaussian') + xlim(0,1)

@@

```{r}
plist = lapply(train.resids, function(df) {
  lim = c(0, max(df$varMAD_median))
  if (max(df$varMAD_median) > 0.05) lim = c(min(df$varMAD_median), 0.05)
  ggplot(df, aes(varMAD_median)) + geom_histogram(color='lightblue') + labs(title=unique(df$kb)) + xlim(lim) + theme_minimal()
})
do.call(grid.arrange, c(plist))

@@

m = melt(all.resids, id.vars=c('kb','cohort'), measure.vars = 'varMAD_median')
ggplot(m %>% filter(value < 0.1), aes(cohort, value)) + facet_grid(~kb) + 
  geom_jitter(width=0.1, color='grey50') + 
  geom_violin(aes(fill=cohort), show.legend = F) +
  geom_boxplot(width=0.1,alpha=0.5) +
  theme_minimal() + 
  labs(title='Post-segmentation MAD residuals', x='median MAD variance')

cutoffs = list('100kb' = 0.009, '15kb' = 0.011, '500kb' = 0.005, '50kb' = 0.011)
sapply(names(train.resids), function(kb) nrow( train.resids[[kb]] %>% filter(varMAD_median >= cutoffs[kb])))
sapply(names(val.resids), function(kb) nrow( val.resids[[kb]] %>% filter(varMAD_median >= cutoffs[kb])))

#ggplot(do.call(bind_rows,train.resids), aes(kb, varMAD_median)) + geom_boxplot() + ylim(0,0.05)
```

@@

ggplot(m, aes(variable, value, color=cohort, group=cohort)) + ylim(0.95,1.05) + facet_grid(~kb) +  
  geom_jitter(position = position_dodge(width=1)) + 
  geom_violin(position = position_dodge(width=1), alpha=0.5) + 
  theme_minimal()

@@

```{r, warning=F, message=F}
z.sds = lapply(train.tiles, function(df) {
  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
})
